---
title: "Problem Set 4"
author: "Tessie Dong, Derek Li, Andi Liu"
date: Due Jan 26th, 2024
output:
    pdf_document:
        keep_tex: true
        extra_dependencies: ['amsmath']
    html_document: default
colorlinks: true
---


<!-- R library setup -->
```{r, setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
rmarkdown::pandoc_available()
library(tidyverse)
library(utils)
library(systemfit)
```

\begin{center}
{\LARGE Part 1: Describe the Data (10 p)}
\end{center}

1. Fill Table \ref{tab:descriptive-stats}'s columns 5 and 6 using, respectively, the data in \texttt{nswpsid.csv} and in \texttt{nswcps.csv}. \textcolor{Gray}{\textbf{Notes}: You want to limit attention to observations with \texttt{treat=0}. You filled columns 3 and 4 in PSet 3.}
    ```{r, message=FALSE, results=FALSE}
    # Load data
    nswpsid <- read_csv("starter-files/nswpsid.csv")
    nswcps <- read_csv("starter-files/nswcps.csv")
    nswpsid_treat0 <- nswpsid %>% filter(treat == 0)
    nswcps_treat0 <- nswcps %>% filter(treat == 0)
    ```
    ```{r}
    summary_cps <- summarise_all(nswcps_treat0, list(mean))
    summary_psid <- summarise_all(nswpsid_treat0, list(mean))
    ```
    <!-- Table imported from `figures/table_descriptive_stats.tex` -->
    \input{figures/table_descriptive_stats.tex}
    

2. Briefly comment on the completed Table \ref{tab:descriptive-stats}.  \textcolor{gray}{\textbf{Hint}: Are the PSID-1 and CPS-1 samples "good" control groups?}

    I would argue that these samples are not the best control groups - this is mostly because many of the OPV covariates from the PSID and CPS exhibit large differences from the characteristics of the NSW sample. For example, the average age of the NSW sample is 25.82, while the average age of the PSID sample is 34.5, and there are large differences in income across the three samples. This suggests that the populations from which PSID and CPS were drawn are not very similar to the population of the NSW sample - making comparisons between treated individuals in the NSW sample and "untreated" individuals in the PSID and CPS samples less reliable, in our opinion.

3. Why do you think that Dehajia and Wahba constructed their "observational datasets" by pulling together the treated sample from NSW and a sample of individuals drawn from either the PSID or the CPS data? \textcolor{gray}{\textbf{Hint:} Both PSID and CPS include information on whether an individual enrolled in a training course during the previous 12 months. Thus, Dehajia and Wahba could have exploited exclusively observational variation in whether an individual enrolled in a training program. Why do you think that they chose not to follow this approach?}

    We believe that Dehajia and Wahba chose to pool the NSW and PSID/CPS datasets because they wanted to have a larger sample size to work with. This is because the NSW sample is relatively small, and the PSID/CPS samples are much larger. By pooling the NSW and PSID/CPS samples, Dehajia and Wahba are able to increase the sample size of their dataset. In addition, by analyzing samples drawn from different distributions (i.e. PSID/CPS datasets), they could increase the generalizability of their results to the population.
    

