---
title: "Problem Set 4"
author: "Tessie Dong, Derek Li, Andi Liu"
date: Due Jan 26th, 2024
output:
    pdf_document:
        keep_tex: true
        extra_dependencies: ['amsmath']
    html_document: default
colorlinks: true
---


<!-- R library setup -->
```{r, setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
rmarkdown::pandoc_available()
library(tidyverse)
library(utils)
library(systemfit)
```

\begin{center}
{\LARGE Part 1: Describe the Data (10 p)}
\end{center}

1. Fill Table \ref{tab:descriptive-stats}'s columns 5 and 6 using, respectively, the data in \texttt{nswpsid.csv} and in \texttt{nswcps.csv}. \textcolor{Gray}{\textbf{Notes}: You want to limit attention to observations with \texttt{treat=0}. You filled columns 3 and 4 in PSet 3.}
    ```{r, message=FALSE, results=FALSE}
    # Load data
    nswpsid <- read_csv("starter-files/nswpsid.csv")
    nswcps <- read_csv("starter-files/nswcps.csv")
    nswpsid_treat0 <- nswpsid %>% filter(treat == 0)
    nswcps_treat0 <- nswcps %>% filter(treat == 0)
    ```
    ```{r}
    summary_cps <- nswcps_treat0 %>% dplyr::summarise_all(list(mean))
    summary_psid <- nswpsid_treat0 %>% dplyr::summarise_all(list(mean))
    ```
    <!-- Table imported from `figures/table_descriptive_stats.tex` -->
    \input{figures/table_descriptive_stats.tex}
    

2. Briefly comment on the completed Table \ref{tab:descriptive-stats}.  \textcolor{gray}{\textbf{Hint}: Are the PSID-1 and CPS-1 samples ``good'' control groups?}

3. Why do you think that Dehajia and Wahba constructed their ``observational datasets'' by pulling together the treated sample from NSW and a sample of individuals drawn from either the PSID or the CPS data? \textcolor{gray}{\textbf{Hint:} Both PSID and CPS include information on whether an individual enrolled in a training course during the previous 12 months. Thus, Dehajia and Wahba could have exploited exclusively observational variation in whether an individual enrolled in a training program. Why do you think that they chose not to follow this approach?}
\end{enumerate}
